Forecasting the black Sigatoka development rate:\\ A comparison of machine learning techniques

### 2016-05-22 Paper Sigatoka - 2015.tex
Some comments to the introduction and SVR

\newcommand{\instead}[2]{\textcolor{dkred}{\st{#1}}#2}

@@

\journal{Journal of \LaTeX\ Templates}

@@

This disease attacks the plant leaves producing a rapid deterioration
of the leaf area, affects the growth and productivity of the plants
due to the impairment of the photosynthetic process, causes a
reduction in the quality of the fruit, and promotes premature
ripening of bunches, which is
the major cause of product losses associated with the black
Sigatoka. \figurename~\ref{fig:diseasestages} shows three stages of
this disease.

@@

This bilogical warning system is based on weekly quantified
observations of the disease progression on young leaves of the plant,
according to Fouré's scale of the symptom stages.
%
The numeric coefficients describe the degree of incidence and the
severity of the disease development.  These coefficients are used to
calculate two variables: gross sum and state of evolution.

The gross sum is based on the present disease progression stage and
the numeric coefficients, which increases with the progression of
the symptoms and the juvenility of the leaf.
%
The state of evolution is calculated using the gross sum and the
foliar emission period.
%
Although threshold levels were initially suggested as a guide to spray
schedules, the fluctuation of these two variables
was found to better define appropiate times to spray
\citep{Marinetal2003}.

@@

The outline of the paper is as follows: in Section~\ref{sec:techs}, we
introduce the machine learning techniques used in this research and
the main related works. Section~\ref{sec:data} summarizes the data the
methodology used in this study.  The results and their discussion are
presented in section~\ref{sec:results}.  We present the main
conclusions and lines for future works in Section~\ref{sec:concl}.

@@

In this section we summary the machine learning techniques used and we present the main related works.

@@

From the perspective of Support Vector Regression (SVR) the regression function $y = f(s)$ for a given dataset $D=\{(s_i,y_i)\}_{i=1}^n$ , is represented as a linear function of the form \citep{Wei2013}:
$$f(s)=w^Ts+b$$
where $w$ and $b$ are respectively the weight vector and the intercept of the model, and they are selected to find an optimal fit to the data available in $D$.

For nonlinear cases, one proceeds by mapping the input p-dimensional vectors via a nonlinear function $\phi : R^p\rightarrow F$, onto the feature space $F$.  After nonlinear mapping, the regression function evolves to a pervasive form:

$$f(s)=w^T \phi (s)+b$$

SVR uses the $\epsilon$-insensitive loss function:
\begin{equation*}
l= {\left| y - f(s) \right| }_{\epsilon } = 
\begin{cases}
   0 & {\left| y - f(s) \right| } \leq \epsilon \\
   \left| y - f(s) \right| - \epsilon & \text{otherwise}
\end{cases}
\end{equation*}
which ignores the error if the difference between the prediction value and the actual value is smaller than $\epsilon$.
%
The $\epsilon$-insensitive loss function allows to find the coefficients $w$ and $b$ by solving a convex optimization problem, which balances the empirical error and the generalization ability.
%
In SVR, the empirical error is measured by the loss function $\epsilon$-insensitive and the generalization ability is measured by the Euclidean norm of $w$ \citep{Cristiani2005}.
%
Then, the optimization problem to identify the regression model can be
formulated by \citep{Wei2013}:

\begin{equation} 
\begin{aligned}
& \underset{}{\text{minimize}}
& & J(w,\xi_i , \xi_i^* ) = \frac{1}{2}   \Bigr| \Bigr| w \Bigr| \Bigr|^2 + C \sum_{i=1}^{n} (\xi_i , \xi_i^* )    \\
& \text{subject to}
& & \begin{array}{lcl} 
y_i - w^T \phi (s) - b  \leq \epsilon + \xi_i   \\
w^T \phi (s) + b - y_i \leq \epsilon + \xi_i^* & i= 1,2,...,n \\
\xi_i , \xi_i^* \geq 0  \\
\end{array}
\end{aligned}
\end{equation}  
where $C$ denotes the penalty parameter between empirical and generalization errors, and  $\xi_i , \xi_i^*$ are slack variables. \figurename
$.$\ref{figura2} shows this situation.

@@

The solution of this optimization problem by the Lagrange method is given by:\\
$$ f(s) = w^T \phi (s) + b = \sum_{i=1}^{n} (\alpha_i - \alpha_i^*) K (s,s_i) + b $$ 
where $\alpha_i - \alpha_i^*$  are the Lagrange multipliers of the optimization problem’s dual form and $K(s_i,s_j )$ is the kernel function satisfying the Mercer condition, and holds:
$$K(s_i,s_j ) = \big \langle  \phi(s_i) , \phi(s_j)  \big \rangle $$
Operations in the kernel function $K(s,s_i )$ are performed in the input space rather than in the potentially high dimensional feature space of $\phi$ \citep{Alonso2013}.

\subsection{Ordinary least squares regression}
This method fits a linear model with coefficients $w = (w1,..,wp)$ to minimize the residual sum of squares between the observed responses in the dataset, and the responses predicted by the linear approximation. Mathematically it solves a problem of the form \citep{scikitlearn2011}:
$$\min_{w} \Bigr| \Bigr| Xw - y \Bigr| \Bigr|_2^2  $$
where $X$ denotes the features matriz. \\
According Pedregosa et al. \citep{scikitlearn2011}  the coefficient estimates for Ordinary Least Squares rely on the independence of the model terms. When terms are correlated and the columns of the design matrix $X$ have an approximate linear dependence, the design matrix becomes close to singular and as a result, the least-squares estimate becomes highly sensitive to random errors in the observed response, producing a large variance. This situation of multicollinearity can arise, for example, when data are collected without an experimental design

@@

\subsection{Echo State Networks (ESN)}
Recurrent Neural Networks (RNN) are useful for temporal patterns, but when they are trained with backpropagation methods, they are very slow.  Echo State Network (ESN) is an alternative training method to solve that problem.  ESN is based on the observation that if a random RNN possesses certain algebraic properties, training only a linear readout from it is often sufficient to achieve excellent performance in practical applications \citep{Lukose2009}. 
For a given training input signal $u(n)  \in R^{N_u}$ a desired target output signal $y^{target}(n) \in R^{N_y}$
is known. Here $n = 1, . . . ,T$ is the discrete time and $T$ is the number of data points in the training dataset. The task is to learn a model with output $y(n) \in R^{N_y}$, where $y(n)$ matches $y^target(n)$ as well as possible, minimizing an error measure $E(y,y^target)$, and, more importantly, generalizes well to unseen data. The untrained RNN part of an ESN is called a dynamical reservoir, and the resulting states x(n) are termed echoes of its input history \citep{Lukose2012}. Finally, these signals are sent to an output layer as shown in the \figurename
$.$\ref{figura3}.
\begin{figure}[h] 
 \centering
 \includegraphics[scale=.9]{Reservorio}
 \caption{An echo state network \citep{Lukose2012}.} 
 \label{figura3} 
\end{figure}

@@

\figurename $.$\ref{figura4} shows the best $R^2$ for each algorithm in the experiment. Results are group by farm. Though La Rita obtains different results in magnitude than 28 Millas, the trend is similar. In both farms, the best results are for linear models, second position is occupied for Echo State Networks and  SVR with gausian and sigmoid kernels are the worst results. In linear models, to predict one week ahead is better than two weeks ahead, and this is better than three weeks ahead.

@@

\figurename $.$\ref{figura5} presents, for one, two and three weeks ahead, the best $R^2$. Results are group by farm. In general, to predict one week ahead is better than two weeks ahead and so on. The number of weeks consider in the observation range in the pattern is not the main discriminant factor, but it is clear that we get better $R^2$ for one week ahead than two weeks ahead and so on.

@@

\figurename $.$\ref{figura6} shows the best $R^2$ for each variables combination. Results are group by farm. The better results are obtained with $\overline{T}_{a}$ and the combination of $\overline{T}_{a}$ with $\overline{W}$, in both farms of similarly. You can note that the use of all variables in the model or the inclusion of the four variables suggest for expert criteria do not improve significantly the results, then the use of more sensors do not assure a better result.

@@

\figurename $.$\ref{figura7} shows the Pareto frontier for each farm with respect to $R^2$ and $RMSE$. The Rita obtains upper $R^2$ with respect to 28 Millas, but 28 Millas obtains better $RMSE$ than La Rita. This situation arise because $RMSE$ considers errors only with respect the prediction and in 28 Millas the average of Stage of Evolution is $4316.16$, unlike, in La Rita the average is $5507.30$. So, in La Rita we obtains higher errors in absolute values. $R2$ is a relative metric between 0 thru 1 and it is less sensitive to absolute values.

@@

\figurename $.$\ref{figura8} shows the best $R^2$ for each algorithm in the experiment. Results are group by farm. Even if linear models continue with good $R^2$, Echo State Networks improve their scores because in $50$ and $100$ last weeks validation, we are in presence of climate change, then the behaviour is less lineal.

@@

\figurename $.$\ref{figura9} presents, for one, two and three weeks ahead, the best $R^2$. Results are group by farm. The results confirms that, in general, to predict one week ahead is better than two weeks ahead, two than three and so on, this for both farms.

@@

\figurename $.$\ref{figura10} shows the best $R^2$ for each variables combination. Results are group by farm. This results confirm that $\overline{T}_{a}$ and the combination of $\overline{T}_{a}$ with $\overline{W}$, in both farms are the best variables combinations.

@@

\figurename $.$\ref{figura11} shows the Pareto frontier for each farm with respect to $R^2$ and $RMSE$.  You can note that the behaviour of $R^2$ and $RMSE$ is similar to the phase one.

>>>>

\newcommand{\instead}[2]{\textcolor{dkred}{\st{#1}}#2}
\newcommand{\figref}[1]{\figurename~\ref{#1}}

@@

\journal{Computers and Electronics in Agriculture}

@@

This disease attacks the plant leaves producing a rapid deterioration
of the leaf area, affects the growth and productivity of the plants
due to the impairment of the photosynthetic process, causes a
reduction in the quality of the fruit, and promotes premature
ripening of bunches, which is
the major cause of product losses associated with the black
Sigatoka. \figref{fig:diseasestages} shows three stages of
this disease.

@@

This bilogical warning system is based on weekly quantified
observations of the disease progression on young leaves of the plant,
according to Fouré's scale of the symptom stages.
%
Numeric coefficients are used to describe the degree of incidence and
the severity of the disease development.  These coefficients are used
to calculate two variables: gross sum and state of evolution.

The gross sum is based on the present disease progression stage and
the numeric coefficients, which increase with the progression of
the symptoms and the juvenility of the leaf.
%
The state of evolution is calculated using the gross sum and the
foliar emission period.
%
Although threshold levels were initially suggested as a guide to spray
schedules, the fluctuation of these two variables
was found to better define appropiate times to spray
\citep{Marinetal2003}.

\TODO{No queda claro threshold levels sobre qué variables!?  Tampoco
  queda claro que quiere decir 'fluctuation' como medida de decisión }

@@

\TODO{Definitivamente ese no es el aporte principal del artículo.
  Cuando esté listo el resto hay que volver aquí.  Creo que va a ser
  algo como mostrar la relevancia de la etapa de preprocesamiento, o
  mostrar que a pesar de ser sistemas complejos, con variables
  caóticas, lo mejor en este caso son predictores lineales, o algo por
  el estilo.}

The outline of the paper is as follows: Section~\ref{sec:techs}
summarizes the machine learning techniques used in this research and
presents the main related works. In Section~\ref{sec:data} we present
the methodology used in this study and describe data used for its
verification.  The results and their discussion are presented in
section~\ref{sec:results}.  The Section~\ref{sec:concl} concludes this
article and presents lines for future works.

@@

In this section we summarize the machine learning techniques used and we present the main related works.

@@

\TODO{Estoy seguro que los lectores van a solicitar reducir esto y
  dejar solo la referencia.  No hay nada nuevo en el resto del paper
  que requiera toda esta información, así que hay que resumirlo a nada
  más lo esencial. }

From the perspective of Support Vector Regression (SVR) the regression
function $y = f(s)$ for a given dataset $D=\{(s_i,y_i)\}_{i=1}^n$ is
represented as a linear function of the form \citep{Wei2013}:
\begin{equation*}
  f(s)=w^Ts+b
\end{equation*}
where $w$ and $b$ are respectively the weight vector and the intercept
of the model, and they are selected to find an optimal fit to the data
available in $D$ in the sense of ...?
\TODO{optimal respecto a qué?.  Hay muchísimas posibilidades de optimalidad!}

For nonlinear cases, the $p$-dimensional input vectors are mapped by a
nonlinear function $\phi : R^p\rightarrow F$ onto a feature space
$F$.  After the nonlinear mapping, the regression function evolves to a
pervasive form:
\begin{equation*}
f(s)=w^T \phi (s)+b
\end{equation*}

In SVR the generalization ability is measured by the Euclidean norm of
$w$, and the empirical error is measured by the $\epsilon$-insensitive
loss function, given by \citep{Cristiani2005}

\TODO{si $l$ es una función, es función de qué, de $s$? $l(s)$? o de
  $y$? o de todos $l(y,s;\epsilon)$?}

\begin{equation*}
l= {\left| y - f(s) \right| }_{\epsilon } = 
\begin{cases}
   0 & {\left| y - f(s) \right| } \leq \epsilon \\
   \left| y - f(s) \right| - \epsilon & \text{otherwise}
\end{cases}
\end{equation*}
which ignores the error if the difference between the prediction value
and the actual value is smaller than $\epsilon$.
%
The $\epsilon$-insensitive loss function allows to find the
coefficients $w$ and $b$ by solving a convex optimization problem,
which balances the empirical error and the generalization ability.
%
Then, the optimization problem to identify the regression model can be
formulated as \citep{Wei2013}:
%
\begin{equation} 
\begin{aligned}
& \underset{}{\text{minimize}}
& & J(w,\xi_i , \xi_i^* ) = \frac{1}{2}   \Bigr| \Bigr| w \Bigr| \Bigr|^2 + C \sum_{i=1}^{n} (\xi_i , \xi_i^* )    \\
& \text{subject to}
& & \begin{array}{lcl} 
y_i - w^T \phi (s) - b  \leq \epsilon + \xi_i   \\
w^T \phi (s) + b - y_i \leq \epsilon + \xi_i^* & i= 1,2,...,n \\
\xi_i , \xi_i^* \geq 0  \\
\end{array}
\end{aligned}
\end{equation}  
where $C$ denotes the penalty parameter between empirical and
generalization errors, and $\xi_i , \xi_i^*$ are slack
variables. \figref{figura2} shows this situation.

@@

The solution of this optimization problem by the Lagrange method is given by:
\begin{equation*}
f(s) = w^T \phi (s) + b = \sum_{i=1}^{n} (\alpha_i - \alpha_i^*) K (s,s_i) + b
\end{equation*}
where $\alpha_i - \alpha_i^*$ are the Lagrange multipliers of the
optimization problem’s dual form and $K(s_i,s_j )$ is the kernel
function satisfying the Mercer condition, and holds:
\begin{equation*}
K(s_i,s_j ) = \big \langle  \phi(s_i) , \phi(s_j)  \big \rangle
\end{equation*}
Operations in the kernel function $K(s,s_i )$ are performed in the
input space rather than in the potentially high dimensional feature
space of $\phi$ \citep{Alonso2013}.

\subsection{Ordinary least squares regression}

This method fits a linear model with coefficients $w = (w_1,..,w_p)$
to minimize the residual sum of squares between the observed responses
in the dataset, and the responses predicted by the linear
approximation. Mathematically it solves a problem of the form
\citep{scikitlearn2011}:
$$\min_{w} \Bigr| \Bigr| Xw - y \Bigr| \Bigr|_2^2  $$
where $X$ denotes the features matrix.

According Pedregosa et al. \citep{scikitlearn2011}  the coefficient estimates for Ordinary Least Squares rely on the independence of the model terms. When terms are correlated and the columns of the design matrix $X$ have an approximate linear dependence, the design matrix becomes close to singular and as a result, the least-squares estimate becomes highly sensitive to random errors in the observed response, producing a large variance. This situation of multicollinearity can arise, for example, when data are collected without an experimental design.

@@

\subsection{Echo State Networks (ESN)}
Recurrent Neural Networks (RNN) are useful for temporal patterns, but when they are trained with backpropagation methods, they are very slow.  Echo State Network (ESN) is an alternative training method to solve that problem.  ESN is based on the observation that if a random RNN possesses certain algebraic properties, training only a linear readout from it is often sufficient to achieve excellent performance in practical applications \citep{Lukose2009}. 
For a given training input signal $u(n)  \in R^{N_u}$ a desired target output signal $y^{target}(n) \in R^{N_y}$
is known. Here $n = 1, . . . ,T$ is the discrete time and $T$ is the number of data points in the training dataset. The task is to learn a model with output $y(n) \in R^{N_y}$, where $y(n)$ matches $y^target(n)$ as well as possible, minimizing an error measure $E(y,y^target)$, and, more importantly, generalizes well to unseen data. The untrained RNN part of an ESN is called a dynamical reservoir, and the resulting states x(n) are termed echoes of its input history \citep{Lukose2012}. Finally, these signals are sent to an output layer as shown in the \figref{figura3}.
\begin{figure}[h] 
 \centering
 \includegraphics[scale=.9]{Reservorio}
 \caption{An echo state network \citep{Lukose2012}.} 
 \label{figura3} 
\end{figure}

@@

\figref{figura4} shows the best $R^2$ for each algorithm in the experiment. Results are group by farm. Though La Rita obtains different results in magnitude than 28 Millas, the trend is similar. In both farms, the best results are for linear models, second position is occupied for Echo State Networks and  SVR with gausian and sigmoid kernels are the worst results. In linear models, to predict one week ahead is better than two weeks ahead, and this is better than three weeks ahead.

@@

\figref{figura5} presents, for one, two and three weeks ahead, the best $R^2$. Results are group by farm. In general, to predict one week ahead is better than two weeks ahead and so on. The number of weeks consider in the observation range in the pattern is not the main discriminant factor, but it is clear that we get better $R^2$ for one week ahead than two weeks ahead and so on.

@@

\figref{figura6} shows the best $R^2$ for each variables combination. Results are group by farm. The better results are obtained with $\overline{T}_{a}$ and the combination of $\overline{T}_{a}$ with $\overline{W}$, in both farms of similarly. You can note that the use of all variables in the model or the inclusion of the four variables suggest for expert criteria do not improve significantly the results, then the use of more sensors do not assure a better result.

@@

\figref{figura7} shows the Pareto frontier for each farm with respect to $R^2$ and $RMSE$. The Rita obtains upper $R^2$ with respect to 28 Millas, but 28 Millas obtains better $RMSE$ than La Rita. This situation arise because $RMSE$ considers errors only with respect the prediction and in 28 Millas the average of Stage of Evolution is $4316.16$, unlike, in La Rita the average is $5507.30$. So, in La Rita we obtains higher errors in absolute values. $R2$ is a relative metric between 0 thru 1 and it is less sensitive to absolute values.

@@

\figref{figura8} shows the best $R^2$ for each algorithm in the experiment. Results are group by farm. Even if linear models continue with good $R^2$, Echo State Networks improve their scores because in $50$ and $100$ last weeks validation, we are in presence of climate change, then the behaviour is less lineal.

@@

\figref{figura9} presents, for one, two and three weeks ahead, the best $R^2$. Results are group by farm. The results confirms that, in general, to predict one week ahead is better than two weeks ahead, two than three and so on, this for both farms.

@@

\figref{figura10} shows the best $R^2$ for each variables combination. Results are group by farm. This results confirm that $\overline{T}_{a}$ and the combination of $\overline{T}_{a}$ with $\overline{W}$, in both farms are the best variables combinations.

@@

\figref{figura11} shows the Pareto frontier for each farm with respect to $R^2$ and $RMSE$.  You can note that the behaviour of $R^2$ and $RMSE$ is similar to the phase one.
